RProtoBuf: Efficient Cross-Language Data Serialization in R

### 2014-01-04 papers/rjournal/eddelbuettel-stokely.Rnw
increment minor version; small update re RProtoBufUtils
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@709 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Saptarshi Guha wrote the RHIPE package \citep{rhipe} which includes
protocol buffer integration with R.  However, this implementation
takes a different approach: any R object is serialized into a message
based on a single catch-all \texttt{proto} schema.  Jeroen Ooms took a
similar approach influenced by Saptarshi in his \pkg{RProtoBufUtils}
package.  Unlike Saptarshi's package, however, RProtoBufUtils depends
on RProtoBuf for underlying message operations.  This package is
available at \url{https://github.com/jeroenooms/RProtoBufUtils}.

The \textbf{RProtoBufUtils} package by Jereoen Ooms provides a
\texttt{serialize\_pb} method to convert R objects into serialized
protocol buffers in this format, and the \texttt{can\_serialize\_pb}
method can be used to determine whether the given R object can safely
be expressed in this way.  To show how how this method works, we
attempt to convert all of the built-in datasets from R into this
serialized protocol buffer representation.

<<echo=TRUE>>=
library(RProtoBufUtils)

datasets <- subset(as.data.frame(data()$results), Package=="datasets")
datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
n <- nrow(datasets)
@

>>>>

Saptarshi Guha wrote the RHIPE package \citep{rhipe} which includes
protocol buffer integration with R.  However, this implementation
takes a different approach: any R object is serialized into a message
based on a single catch-all \texttt{proto} schema.  Jeroen Ooms took a
similar approach influenced by Saptarshi in the \pkg{RProtoBufUtils}
package (which has now been integrated in \pkg{RProtoBuf}).  Unlike
Saptarshi's package, however, RProtoBufUtils depends 
on, and extends, RProtoBuf for underlying message operations.

One key extension of \pkg{RProtoBufUtils} is the 
\texttt{serialize\_pb} method to convert R objects into serialized
protocol buffers in the catch-all schema. The \texttt{can\_serialize\_pb}
method can be used to determine whether the given R object can safely
be expressed in this way.  To illustrate how this method works, we
attempt to convert all of the built-in datasets from R into this
serialized protocol buffer representation.

<<echo=TRUE>>=
datasets <- subset(as.data.frame(data()$results), Package=="datasets")
datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
n <- nrow(datasets)
@
